---
title: "Ben Anderson"
author: 'Colour Blind Palettes for Plots (Contact: b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    self_contained: no
    toc: yes
    toc_depth: 2
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
# colour blind palettes for charts
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette
# with grey
cbgPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# with black
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Colour Blind palettes

Obviously these aid interpretation of plots by those who are colour blind but the [palettes](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette) are also useful clear defaults. Note that there are `r length(cbbPalette)` colours in each palette:

 * palette with grey: `r cbgPalette` 
 * palette with black: `r cbbPalette` 

Source: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette

This means that a plot with more than this number of factors will run out of colours (see below).

## Grey palette

This is a cross-reference to Figure \@ref(fig:plot1) which uses the grey palette.

```{r plot1, fig.cap="Scatter plot of mpg by displacement (grey palette)"}
df <- mtcars

df$auto <- ifelse(df$am == 0, "Automatic", "Manual") # set labels
p <- ggplot2::ggplot(df, aes(x = mpg, y = disp, colour = as.factor(carb))) +
  guides(colour = guide_legend(title = "Number of carburetors:")) +
  theme(legend.position="bottom") +
  scale_colour_manual(values=cbgPalette) + # use colour-blind friendly grey palette
  geom_point() # <- make the plot in an object first
p + labs(x = "Miles/(US) gallon", y = "Displacement (cu.in.)") + facet_grid(auto ~ .) # <- draw the plot and add more features
```

## Black palette

This is a cross-reference to Figure \@ref(fig:plot2) which uses the black palette.

```{r plot2, fig.cap="Scatter plot of mpg by displacement (black palette)"}


p <- ggplot2::ggplot(df, aes(x = mpg, y = disp, colour = as.factor(carb))) +
  guides(colour = guide_legend(title = "Number of carburetors:")) +
  theme(legend.position="bottom") +
  scale_colour_manual(values=cbbPalette) + # use colour-blind friendly grey palette
  geom_point() # <- make the plot in an object first
p + labs(x = "Miles/(US) gallon", y = "Displacement (cu.in.)") + facet_grid(auto ~ .) # <- draw the plot and add more features
```

# Using viridis

An alternative to the palettes is the [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) package [@viridis]. This has the benefit of not having a small number of colours although Figure \@ref(fig:plot3) doesn't really demonstrate this!

```{r plot3, fig.cap="Scatter plot of mpg by displacement (viridis palette)"}
library(viridis)
p <- ggplot2::ggplot(df, aes(x = mpg, y = disp, colour = carb)) +
  guides(colour = guide_legend(title = "Number of carburetors:")) +
  theme(legend.position="bottom") +
  scale_colour_viridis() + # use viridis
  geom_point() # <- make the plot in an object first
p + labs(x = "Miles/(US) gallon", y = "Displacement (cu.in.)") + facet_grid(auto ~ .) # <- draw the plot and add more features
```

# Summary

Personally I prefer the [colour blind palettes](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette) if I have a small number of categories (less than `r length(cbbPalette)`) but usually use [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html) if I have more than that and if I have continuous data.

# Ackowledgement
Uses ggplot2 [@ggplot2] and associated mtcars dataset to draw plots. Palettes available from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#a-colorblind-friendly-palette

# References
