---
title: "Testing drake"
author: '@dataknut'
date: "4/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries needed in the .Rmd
library(kableExtra)
```

# Drake

Use the [drake](https://ropenscilabs.github.io/drake-manual/) R package to make data flow, analysis and reporting pipelines. Drake looks after which bits of the pipeline need refreshing every time you re-run the drake _plan_. So when you run the report as part of your plan (see [drake.R](drake.R)), you can `readd` the objects you created in the R script to report them in this .Rmd file. For example, here we `readd` NZ Electricity Authority generation data for 1/1/2018 which we downloaded in drake.R:

```{r cars}
# readd the data object

dt <- readd(data)

t <- summary(dt)

kableExtra::kable(t, caption = "Data summary")
```

Now re `readd` a plot we made using the data. Note that we don't build the plot in the .Rmd file (via knitr) we just bring back the object from wherever drake stored it.

```{r pressure}
p <- readd(profilePlot) # we don't need to create the p but this would let us add stuff

p
```

This is all really rather cool...

But the very cool bit is that if we make some changes to the code that makes the plot, when we re-make our drake plan, drake will only re-build the plot and the report. It will not go and get the data again - it knows nothing changed.
