---
params:
  authors: 'Ben Anderson'
  title: 'Testing gt for tables'
title: '`r paste0(params$title)`'
subtitle: '`r paste0(params$subtitle)`'
author: '`r paste0(params$authors)` (Contact: b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
always_allow_html: yes
output:
  bookdown::html_document2:
    code_folding: hide
    number_sections: yes
    self_contained: no
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
  bookdown::pdf_document2:
    toc: yes
    toc_depth: 2
bibliography: statsCodeRefs.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries needed in the .Rmd

library(tidyverse)
library(glue)

library(data.table) # last to avoid masking

library(gt)

```

# Introduction

`kableExtra` makes great tables in html & pdf but not so good in MS Word. Sadly many of our co-authors still like to use MS Word. So we need a Word-friendly way to make nice tables...

# gt

The [gt package](https://github.com/rstudio/gt) might help.

```{r getData}
df <- gtcars
dt <- as.data.table(df) # cos we like data.tables

t <- table(dt$mfr, dt$bdy_style)

gt(t)
```

So a gt object is clearly in long form. The examples all use pipes to pass the table through dplyr stuff...

```{r test}


# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  dplyr::filter(date >= start_date & date <= end_date) %>%
  dplyr::select(-adj_close) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue::glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = vars(date),
    date_style = 3
  ) %>%
  fmt_currency(
    columns = vars(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = vars(volume),
    suffixing = TRUE
  )
```

Well the html version works nicely. The Word version does not <sigh>. We really don't want to have to export to rtf and then paste back in...

# R environment

Packages used:

 * gt [@gt]
 
And:
 
 * tidyverse - [@tidyverse]
 * glue - [@glue]
 * data.table - [@data.table]
 
# References
