---
title: "Testing the survey package"
author: Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)
date: 'Last run at: `r Sys.time()`'
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
bibliography: ~/bibliography.bib
---

# Libraries used

 * survey [@survey] - for survey data (weighting etc)
 * ggplot [@ggplot2] - for graphs
 * data.table [@data.table] - for fast big tables
 * skimr [@skimr] - for summaries
 * knitr [@knitr] - for knited tables


```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(survey)
library(ggplot2)
library(data.table)
library(skimr)
library(knitr)
```

Load some data that looks a bit like survey data (ggplot's diamonds data) and turn it into a data.table. Then create a random value between 0 and 1 which we will use as a pretend survey non-response weight.

```{r set data}
diamondsDT <- data.table(diamonds)
skim(diamonds)

diamondsDT <- diamondsDT[, wt := runif(nrow(diamondsDT))]

t <- diamondsDT[, .(Mean = mean(wt), Min = min(wt), Max = max(wt), N = .N)]

kable(caption = "Distribution of wt", t)

# create a fake unique 'id' for each row
diamondsDT <- diamondsDT[, id := seq(1:nrow(diamondsDT))]

# create 1/0 code for cut = Fair
diamondsDT <- diamondsDT[, isFair := ifelse(cut == "Fair", 1, 0)]

# check
with(diamondsDT, table(isFair, cut, useNA = "always"))
```

Now we use the wt to pretend to create a large survey design object which has a non-response weight = wt.

```{r Set weight}
diamondsSVY <- svydesign(ids = ~ id, 
                                    weights = ~wt, 
                                    data = diamondsDT
)
```

Now compare a simple and a weighted table.

```{r Test simple tables}

uwt <- table(diamondsDT$cut, diamondsDT$color)

kable(caption = "Cut vs colour, unweighted", uwt)

wt <- svytable(~ cut + color, diamondsSVY)

kable(caption = "Cut vs colour, unweighted", round(wt,2))
  
```

Something of a difference...

Now test what happens when we set a category to 1/0 and take the mean and proportion.

```{r Test survey mean and proportions}

meanT <- svymean(isFair ~ color, diamondsSVY)

print("Using svymean")
meanT
confint(meanT)

propT <- svyciprop(~ isFair, diamondsSVY)

print("Using svyciprop")
propT

tableT <- svytable(~ color + isFair, diamondsSVY)
print("Using svytable")
tableT

tableDT <- as.data.table(tableT)


```


# References