---
title: "Nomisr testing"
author: 'Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)'
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    code_folding: hide
    fig_caption: true
    keep_md: true
    number_sections: true
    self_contained: no
    toc: true
    toc_float: true
    toc_depth: 2
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(nomisr)
library(data.table) # for big data
library(knitr) # for kable etc
```

# nomisr

An R interface to UK Census data via [nomis](https://www.nomisweb.co.uk/api/v01/help). For more info see:

 * https://cran.r-project.org/web/packages/nomisr/index.html
 * https://github.com/ropensci/nomisr
 * Citation - [@nomisr]
 
This code also uses data.table [@data.table].

# Get full list of tables

nomisr provides various ways to search the tables list but it is a bit easier on the API to get the list as a data table and then search it locally...

```{r getTableList}

nomisTablesDT <- as.data.table(nomis_data_info())

knitr::kable(caption = paste0("List of first 5 NOMIS tables (of ", nrow(nomisTablesDT)), head(nomisTablesDT, 5))
```

That gives a list of `r nrow(nomisTablesDT)` tables.

# Tables on heating

The following are the tables that have 'heating' in the name...

```{r getHeatingTableList}

knitr::kable(caption = "List of NOMIS tables with 'heating' in the name", nomisTablesDT[name.value %like% "heat"])
```

Let's see what geographies are available for 'NM_146_1'.

```{r getHeatingTableGeoList}

knitr::kable(caption = "NM_146_1 geography list", 
             nomis_get_metadata(id = "NM_146_1", concept = "geography", type = "type"))
```

So we can get OAs & LSOAs etc.

Get the data at local authority level... lower levels of geography generate time-outs on the API if we do not have a key.


```{r getHeatingTableData}

NM_146_1_DT <- as.data.table(nomis_get_data(id = "NM_146_1", time = "latest", geography = "TYPE464"))

knitr::kable(caption = paste0("NM_146_1 Local Authority level (first 5 rows of ", nrow(NM_146_1_DT)),
             head(NM_146_1_DT,5))
```



# References

