---
title: "OLS Regression modeling in R"
author: Ben Anderson (b.anderson@soton.ac.uk, `@dataknut`)
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    keep_md: no
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: yes
  word_document:
    fig_caption: yes
    toc: yes
    toc_depth: 4
bibliography: '`r path.expand("~/bibliography.bib")`'
---
```{r knitrSetUp, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # echo code so reader can see what is happening
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig_caption = TRUE)
knitr::opts_chunk$set(fig_height = 6) # default, make it bigger to stretch vertical axis
knitr::opts_chunk$set(fig_width = 8) # full width
knitr::opts_chunk$set(tidy = TRUE) # tidy up code in case echo = TRUE
```

```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- Sys.time()

# Load required packages ----
reqLibs <- c("rms",
       "stargazer",
       "car",
       "broom",
       "ggplot2",
       "data.table"
       )
       
# do this to install them first if needed
#install.packages(x)
print("Loading required packages")

# be careful - this will return a FALSE if a package doesn't load but the script will NOT stop!
lapply(reqLibs, require, character.only = T)
```

# Why are we here?

To learn how to do ols regression modelling in R (markdown) and specifically how to use broom [@broom] and car [@car] for diagnostics, stargazer [@stargazer] for model tables, data.table [@data.table] for data manipulation and ggplot [@ggplot2] for results visualisation. 

For more on regression try: http://socserv.socsci.mcmaster.ca/jfox/Courses/Brazil-2009/index.html 

So many birds and just one stone...

```{r set data}
# R has a very useful built-in dataset called mtcars
# http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

# A data frame with 32 observations on 11 variables.
# [, 1] 	mpg 	Miles/(US) gallon
# [, 2] 	cyl 	Number of cylinders
# [, 3] 	disp 	Displacement (cu.in.)
# [, 4] 	hp 	Gross horsepower
# [, 5] 	drat 	Rear axle ratio
# [, 6] 	wt 	Weight (1000 lbs)
# [, 7] 	qsec 	1/4 mile time
# [, 8] 	vs 	V/S
# [, 9] 	am 	Transmission (0 = automatic, 1 = manual)
# [,10] 	gear 	Number of forward gears
# [,11] 	carb 	Number of carburetors 


# Load mtcars ----
mtcars <- mtcars

summary(mtcars) # base method
```

# Examine dataset

```{r examine data}
names(mtcars)
pairs(~mpg+disp+hp+drat+wt,labels=c(
  "Mpg","Displacement","Horse power", 
  "Rear axle rotation","Weight"), data=mtcars, main="Simple Scatterplot Matrix")
```

Test normality of mpg (outcome variable of interest) as linear models rest on this assumption (and quite a few others...).

```{r normality}
# test with a histogram
hist(mtcars$mpg)

# test with a qq plot
qqnorm(mtcars$mpg); qqline(mtcars$mpg, col = 2)

shapiro.test(mtcars$mpg)
# if p > 0.05 => normal 

# is it? Beware: shapiro-wilks is less robust as N -> 
```

Mpg seems to approximate normality but with a slightly worrying tail effect at the right hand extreme.

# Model with 1 term predicting mpg

Run a model trying to see if qsec predicts mpg. This will print out a basic table of results.

```{r model.1}
# qsec = time to go 1/4 mile from stationary
mpgModel1 <- lm(mpg ~ qsec, mtcars)

# results?
summary(mpgModel1)
```

Now run the standard diagnostics over the model results.

```{r model.1.diag}
# Diagnostics ----

message("# Diagnostic plots")

plot(mpgModel1)

message("# Normality of residuals")

hist(mpgModel1$residuals)

qqnorm(mpgModel1$residuals); qqline(mpgModel1$residuals, col = 2)

shapiro.test(mpgModel1$residuals)

message("# Normality of standardised residuals")

# it is usual to do these checks for standardised residuals - but the results are the same
# add casewise diagnostics back into dataframe
mtcars$studentised.residuals <- rstudent(mpgModel1)

qqnorm(mtcars$studentised.residuals); qqline(mtcars$studentised.residuals, col = 2)

shapiro.test(mtcars$studentised.residuals)

# if p > 0.05 => normal 
# is it?
# But don't rely on the test espcially with large n
```

Now try using the car package [@car] to do the same things.

```{r model.1.diag.car}
# The 'car' package has some nice graphs to help here
car::qqPlot(mpgModel1) # shows default 95% CI
car::spreadLevelPlot(mpgModel1)
message("# Do we think the variance of the residuals is constant?")
message("# Did the plot suggest a transformation? If so, why?")

message("# autocorrelation/independence of errors")
car::durbinWatsonTest(mpgModel1)
# if p < 0.05 then a problem as implies autocorrelation
# what should we conclude? Why? Could you have spotted that in the model summary?

message("# homoskedasticity")
plot(mtcars$mpg,mpgModel1$residuals)
abline(h = mean(mpgModel1$residuals), col = "red") # add the mean of the residuals (yay, it's zero!)

message("# homoskedasticity: formal test")
car::ncvTest(mpgModel1)
# if p > 0.05 then there is heteroskedasticity
# what do we conclude from the tests?
```

go back to the model - what can we conclude from it?

```{r report model 1}
summary(mpgModel1)
```

There are an infinite number of ways to report regression model results and every journal has it's own. We look at this in detail below using stargazer [@stargazer] but the [guidance here](https://www.csus.edu/indiv/v/vangaasbeckk/courses/200a/sup/regressionresults.pdf) is also quite useful.

# Model with more than 1 term

So our model was mostly OK (one violated assumption?) but the r sq was quite low. 

Maybe we should add the car's weight?

```{r model.2}
# wt = weight of car
mpgModel2 <- lm(mpg ~ qsec + wt, mtcars)

# results?
summary(mpgModel2)
```

Run diagnostics.

```{r model.2.diag}
# Diagnostics ----

car::qqPlot(mpgModel2) # shows default 95% CI
car::spreadLevelPlot(mpgModel2)
message("# Do we think the variance of the residuals is constant?")
message("# Did the plot suggest a transformation? If so, why?")

message("# autocorrelation/independence of errors")
car::durbinWatsonTest(mpgModel2)
# if p < 0.05 then a problem as implies autocorrelation
# what should we conclude? Why? Could you have spotted that in the model summary?

message("# homoskedasticity")
plot(mtcars$mpg,mpgModel2$residuals)
abline(h = mean(mpgModel2$residuals), col = "red") # add the mean of the residuals (yay, it's zero!)

message("# homoskedasticity: formal test")
car::ncvTest(mpgModel2)
# if p > 0.05 then there is heteroskedasticity

# but also
message("# additional assumption checks (now there are 2 predictors)")

message("# -> collinearity")
car::vif(mpgModel2)
# if any values > 10 -> problem

message("# -> tolerance")
1/car::vif(mpgModel2)
# if any values < 0.2 -> possible problem
# if any values < 0.1 -> definitely a problem

# what do we conclude from the tests?
```

Whilst we're here we should also plot the residuals for model 2 against the fitted values (as opposed to the observed values which we did earlier). h/t to https://gist.github.com/apreshill/9d33891b5f9be4669ada20f76f101baa for this.

```{r model.2.plot.residuals}
# save the residuals via broom
resids <- augment(mpgModel2)

# plot fitted vs residuals
ggplot(resids, aes(x = .fitted, y = .resid)) + 
    geom_point(size = 1)
```

As with a plot of residuals against observed values, hopefully we didn't see any obviously strange patterns (unlike that [gist](https://gist.github.com/apreshill/9d33891b5f9be4669ada20f76f101baa)).

Now compare the models

```{r models.compare}
# comparing models

message("# test significant difference between models")
anova(mpgModel1, mpgModel2)
# what should we conclude from that?
```


# Reporting OLS results with confidence intervals

Reporting should include coefficients, p values and confidence intervals for factors in each model as well as regression diagnostics so that readers can judge the goodness of fit and uncertainty for themselves (see https://amstat.tandfonline.com/doi/full/10.1080/00031305.2016.1154108 for guidance and also https://www.csus.edu/indiv/v/vangaasbeckk/courses/200a/sup/regressionresults.pdf – with the addition of e.g. 95% confidence intervals to the tables).

The p values tell you whether the 'effect' (co-efficient) is statistically significant at a given confidence threshold. By convention this is usually one of p < 0.05 (5%), p < 0.01 (1%), or p < 0.001 (0.1%). Sometimes, this is (justifiably) relaxed to p < 0.1 (10%) for example. The choice of which to use is normative and driven by your [appetite for Type 1 & Type 2 error risks](https://git.soton.ac.uk/ba1e12/weGotThePower) and the uses to which you will put the results.

But only _you_ can decide if it is IMPORTANT!

It is usually best to calculate and inspect confidence intervals for your estimates alongside the p values.

This indicates:

 * statistical significance (if the CIs do not include 0) - just like the p value
 * precision - the width of the CIs shows you how precise your estimate is

You can calculate them using the standard error (s.e.) from the summary:

 * lower = estimate - (s.e.*1.96) 
 * upper = estimate + (s.e.*1.96)
 * just like for t tests etc (in fact this _is_ a t test!!)

Or use `confint()` which is more precise.

Print out the summaries again and calculate 95% confidence intervals and p values each time. 

```{r model.1.add.CI}
# Model 1
# save results as log odds
# the cbind function simply 'glues' the columns together side by side
mpgModel1Results_CI <- cbind(Coef = coef(mpgModel1),
                              confint(mpgModel1),
                             p = round(summary(mpgModel1)$coefficients[,4], 3)
                             )
mpgModel1Results_CI
```

Notice that where p < 0.05 our 95% CI does not include 0. These tell you the same thing: that with a 95% threshold, the coefficient's differnece from 0 is statistically significant. Notice this is not the case for the constant (Intercept).

Now we do that again but for extra practice we use a bonferroni correction to take into account the number of predictors we used. As with most things in statistics, there is active debate about the use of this...

```{r model.1.add.CI.bonf}
# Model 1 bf
# use confint to report confidence intervals with bonferroni corrected level
bc_p1 <- 0.05/length(mpgModel1$coefficients)

# save results as log odds
# the cbind function simply 'glues' the columns together side by side
mpgModel1Results_bf <- cbind(Coef = coef(mpgModel1),
                              confint(mpgModel1, level = 1 - bc_p1),
                             p = round(summary(mpgModel1)$coefficients[,4], 3)
                             )
mpgModel1Results_bf
```

The coefficient has not change (of course) and nor has the default p value produced by the original lm model but our confidence intervals have been adjusted. You can see that we are now using the more stringent 2.5% and the intervals are wider. We would still conclude there is a statistically significant effect at this threshold but we are a bit less certain.

> Is that the right interpretation?

Now do the same for model 2.

```{r model.2.add.CI}
# Model 2 bf
# use confint to report confidence intervals with bonferroni corrected level
bc_p2 <- 0.05/length(mpgModel2$coefficients)

# save results as log odds
# the cbind function simply 'glues' the columns together side by side
mpgModel2Results_bf <- cbind(Coef = coef(mpgModel2),
                             confint(mpgModel2,level = 1 - bc_p2),
                             p = round(summary(mpgModel2)$coefficients[,4], 3)
                             )

mpgModel2Results_bf
```

# Reporting with Stargazer

Now we'll try reporting the two model using the stargazer package [@stargazer] to get pretty tables. Note we use options to automatically create the 95% CI and to report the results on just one line per predictor. This is especially helpful for models with a lot of variables. However you will see that the default is to report p values as *.

```{r use stargazer, results='asis'}
stargazer(mpgModel1, mpgModel2, 
          title = "Model results",
          ci = TRUE, 
          single.row = TRUE,type = "html")
```


We can add the p values using stargazer options. Note that this puts them under the 95% CI - we may want them in a new column.


```{r use.stargazer.add.P, results='asis'}
stargazer(mpgModel1, mpgModel2, 
          title = "Model results",
          ci = TRUE, 
          report = c("vcsp*"),
          single.row = TRUE,type = "html")
```

## Stargazer style options

Stargazer can simulate a range of journal output formats.

Let's start with 'all' which prints everything. This is quite good for your first model report but probably wouldn't go in an article.

```{r use.stargazer.all, results='asis'}
stargazer(mpgModel1, mpgModel2, 
          title = "Model results",
          ci = TRUE, 
          style = "all",
          single.row = TRUE,type = "html")
```

### American Economic Review

Now let's try American Economic Review.

```{r use.stargazer.aer, results='asis'}
stargazer(mpgModel1, mpgModel2, 
          title = "Model results",
          ci = TRUE, 
          style = "aer",
          single.row = TRUE,type = "html")
```

### Demography

Or perhaps Demography.

```{r use.stargazer.demography, results='asis'}
stargazer(mpgModel1, mpgModel2, 
          title = "Model results",
          ci = TRUE, 
          style = "demography",
          single.row = TRUE,type = "html")
```

# Visualise results using ggplot

Now we'll visualise them using broom and ggplot as we've found that non-statisticians can interpret plots more easily. This is especially useful for showing what the 95% confidence intervals 'mean'.

Just to confuse you we're going to convert the data frames produced by broom in to [data.tables](https://github.com/Rdatatable/data.table/wiki) [@data.table]. It makes little difference here (you can use data frames where we use data.tables) but `data.table::` is good to get to know for _data science_ because it does a lot of data wrangling tasks [very fast](https://h2oai.github.io/db-benchmark/).

```{r visualise models}
# Broom converts the model results into a data frame (very useful!)
# We then convert it to a data.table. We like to use the DT suffix to remind us this is a data.table
mpgModel1DT <- as.data.table(tidy(mpgModel1))

# add the 95% CI
mpgModel1DT$ci_lower <- mpgModel1DT$estimate - qnorm(0.975) * mpgModel1DT$std.error 
mpgModel1DT$ci_upper <- mpgModel1DT$estimate + qnorm(0.975) * mpgModel1DT$std.error
mpgModel1DT <- mpgModel1DT[, model := "Model 1"] # add model label for ggplot to pick up

# repeat for model 2
mpgModel2DT <- as.data.table(tidy(mpgModel2))
mpgModel2DT$ci_lower <- mpgModel2DT$estimate - qnorm(0.975) * mpgModel2DT$std.error 
mpgModel2DT$ci_upper <- mpgModel2DT$estimate + qnorm(0.975) * mpgModel2DT$std.error
mpgModel2DT <- mpgModel2DT[, model := "Model 2"] # add model label for ggplot to pick up

# rbind the data tables so you have long form data
modelsDT <- rbind(mpgModel1DT, mpgModel2DT)

# plot the coefficients using colour to distinguish the models and plot the 95% CIs
ggplot(modelsDT, aes(x=term, y=estimate, fill = model)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(aes(ymin=ci_lower, 
                    ymax=ci_upper),
                    width=.2,
                    position=position_dodge(.9)
                    ) +
  labs(title = "Model results",
       x = 'Variable',
       y = 'Coefficient',
       caption = paste0("Model 1 & 2 results, Error bars = 95% CI",
                        "\n Data: mtcars")
       ) +
  coord_flip() # rotate for legibility
                
```

So now we can easily 'see' and interpret our results.

# About

## Runtime

Analysis completed in: `r round(Sys.time() - startTime, 2)` seconds using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

R packages used (`r reqLibs`):

 * base R - for the basics [@baseR]
 * ggplot2 - for slick graphics [@ggplot2]
 * car - for regression diagnostics [@car]
 * broom - for tidy model results [@broom]
 * data.table - for fast data manipulation [@data.table]
 * knitr - to create this document [@knitr]
                     
# References
