---
title: "References in Rmd"
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output: 
  bookdown::html_document2:
    keep_md: no
    code_folding: hide
    collapsed: no
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
subtitle: Notes
bibliography: statsCodeRefs.bib
---

# Citations {#citations}

`[@broom]` -> [@broom] <- single ref

`[@broom; @knitr]` -> [@broom; @knitr] <- multiple refs

This form `[-@broom]` [-@broom] drops author name compared to `[@broom]` [@broom] which doesn't.

Get citation styles from [githb](https://github.com/citation-style-language/styles) & install locally?

See also http://citationstyles.org/ 

# Cross-references

If we use the `bookdown`  package [@bookdown] then we can do [cross-references]() to tables, figures and sections.

Table \@ref(tab:tableKnitr) shows something interesting. 

> NB: This breaks if you have a range of [non-alphanumeric characters](https://bookdown.org/yihui/rmarkdown/bookdown-markdown.html#cross-referencing) in your chunk labels. 

This means we can't use our preferred '.' to seperate chunk label words but there we go...

```{r tableKnitr}
knitr::kable(summary(mtcars), caption = "Table rendered using knitr::kable")
```

Table \@ref(tab:tableKableExtra) shows something interesting with fomrating applied.

```{r tableKableExtra}
library(kableExtra)
kableExtra::kable(summary(mtcars), caption = "Table rendered using kableExtra::kable") %>%
  kable_styling()
```

If we want to hop back to a section we need a cross-refence to it e.g. to Section \@ref(citations).

# References
