---
params:
  title: ""
title: '`r params$title`'
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
bibliography: '`r path.expand("~/bibliography.bib")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(kableExtra)

```

 * One:
    * One.1;
    * One.2.

1. One (the '.' is crucial):
    i) One.1;
    ii) One.2.
2. Two

# Introduction {#intro}

This is Section \@ref(intro)

# Plots {#plots}

See Figure \@ref(fig:myPlot) in Section \@ref(plots)

```{r myPlot, fig.cap="cars mpg plot"}
plot(mtcars$mpg)
```

# Tables {#tables}

See Table \@ref(tab:myKable) in Section \@ref(tables).

```{r myKable}
df <- mtcars[1:5, 1:6]
knitr::kable(df, caption = "mtcars table (knitr::kable)")
```

See Table \@ref(tab:myKableExtra) in Section \@ref(tables).

```{r myKableExtra}
kableExtra::kable(df, caption = "mtcars table (kableExtra::kable)")
```

See Table \@ref(tab:myKableExtraPretty) in Section \@ref(tables).

```{r myKableExtraPretty}
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
# fails on knit to pdf
kableExtra::kable(df, caption = "mtcars table (kableExtra::kable with kable_styling())") %>%
  kable_styling()
```

# R packages

Uses:

 * knitr [@knitr] 
 * kableExtra [@kableExtra] 
 * rmarkdown [@rmarkdown]
 * bookdown [@bookdown]
 
# References