---
title: 'R Notebook: Stargazer Package Tests'
author: Ben Anderson (b.anderson@soton.ac.uk/@dataknut) [Energy & Climate Change,
  Faculty of Engineering & Environment, University of Southampton]
date: 'Last run at: `r Sys.time()`'
output:
  pdf_document:
    toc: yes
    number_sections: yes
  html_document:
    toc: yes
bibliography: ~/bibliography.bib
---

# Setup
```{r setup}
startTime <- Sys.time() # always good to time your code
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = TRUE)
knitr::opts_chunk$set(message = TRUE)
knitr::opts_chunk$set(fig_caption = TRUE)
knitr::opts_chunk$set(tidy = TRUE)


library(knitr) # for kable
library(stargazer)
```

Key packages used:

 * base R - for the basics [@baseR]
 * stargazer - for pretty tables [@stargazer]
 * knitr - to create this document [@knitr]
 
# Summary

We can print an ordinary table:

```{r oldTable}
summary(cars)
```

# Kable

We can also do that using kable:

```{r kableTable}
t <- summary(cars)
kable(caption = "Cars",
      t
      )
```

# Stargazer

We can also use stargazer:
```{r stargazerTable, results='asis'}
# NB: we need to force knitr to keep the html (asis)
stargazer(cars, type = "html")
```

Stargazer is especially good at regression model outputs:

```{r modelTable, results='asis'}
model1 <- lm(mpg ~ disp + cyl, mtcars)
model2 <- lm(mpg ~ disp + cyl + hp + qsec, mtcars)

stargazer(model1, model2, type = "html",
          column.sep.width = "2pt", # default = 5pt
          font.size = "tiny" # tiny, scriptsize, footnotesize,small,normalsize,large
          )
```

Stargazer has a lot of [options](https://www.rdocumentation.org/packages/stargazer/versions/5.2/topics/stargazer) that can be used to change what goes where...

***
__Meta:__
Analysis completed in: `r round(Sys.time() - startTime,3)` seconds using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com).

# References
